---
title: "Visualising COVID-19"
author: "JUSTIN LO"
date: "2023-12-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE, message=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
```

Reading in the dataset of worldwide confirmed cases and confirmed cases of china and non china countries and assigning it to a variable
```{r, echo=FALSE}
confirmed_cases_worldwide<- read_csv("confirmed_cases_worldwide.csv")
head(confirmed_cases_worldwide)

confirmed_cases_china_vs_world <- read_csv("confirmed_cases_china_vs_world.csv")
head(confirmed_cases_china_vs_world)
```


Visualization of confirmed cases trend and confirmed cases of china and non china countries
```{r, out.width='80%'}
plt_cum_confirmed_cases_worldwide<-ggplot(data = confirmed_cases_worldwide, aes(x=date, y=cum_cases))+geom_line()+ylab("Cumulative confirmed cases")+ ggtitle("Covid cases worldwide")
plt_cum_confirmed_cases_worldwide

plt_cum_confirmed_cases_china_vs_world <- ggplot(data=confirmed_cases_china_vs_world, aes(x= date, y= cum_cases, color = is_china)) +
  geom_line() +
  ylab("Cumulative confirmed cases") +
  ggtitle("Covid cases sepearted by China and non china countries")
plt_cum_confirmed_cases_china_vs_world
```



It was then realised that the huge jump in the China line on February 13, 2020 wasn't just a bad day regarding the outbreak; China changed the way it reported figures on that day (CT scans were accepted as evidence for COVID-19, rather than only lab tests).

To reflect this on the plot, we can add x-intercepts and labels to the plot
```{r}
who_events <- tribble(
  ~ date, ~ event,
  "2020-01-30", "Global health\nemergency\ndeclared",
  "2020-03-11", "Pandemic\ndeclared",
  "2020-02-13", "China reporting\nchanged"
) %>%
  mutate(date = as.Date(date))

plt_cum_confirmed_cases_china_vs_world_2<-plt_cum_confirmed_cases_china_vs_world +
  geom_vline(aes(xintercept = date), data = who_events, linetype = "dashed") +
  geom_label(aes(date, label = event), data = who_events, y = 1e5, color="black")
plt_cum_confirmed_cases_china_vs_world_2
```


Now, to investigate china's confirmed cases after feb15. 
I have created a line plot showing the data of china's confirmed cases after feb 15.
```{r}
china_after_feb15<- confirmed_cases_china_vs_world %>% 
  filter(is_china=="China", date>="2020-02-15")
head(china_after_feb15)

plt_cum_confirmed_cases_china_after_feb15<-ggplot(data= china_after_feb15, aes(x=date, y=cum_cases))+geom_line()+ylab("cumulative confirmed cases")+ggtitle("Covid cases in China")
plt_cum_confirmed_cases_china_after_feb15
```


To create the same plots but for non-china countries
```{r}
no_china<- confirmed_cases_china_vs_world %>% 
  filter(is_china!="China")

yes_china<- confirmed_cases_china_vs_world %>% 
  filter(is_china=="China")

plt_not_china_trend_line<- ggplot(data = no_china, aes(x=date, y=cum_cases)) + geom_line() +ylab("cumulative confirmed cases")+ggtitle("Covid Cases in non-china countries")
plt_not_china_trend_line

```


Import data on covid cases by country with China excluded. 
Then, calculate which country has had the highest cumulative cases
```{r}
confirmed_cases_by_country<- read_csv("confirmed_cases_by_country.csv")
head(confirmed_cases_by_country)

top_countries_by_total_cases<-confirmed_cases_by_country %>% 
  group_by(country) %>% 
  summarize(total_cases = max(cum_cases)) %>%
  top_n(7, total_cases)

top_countries_by_total_cases
```


To create line plots of cumulative confirmed cases by country
```{r}
confirmed_cases_top7_outside_china <- read.csv("confirmed_cases_top7_outside_china.csv")
head(confirmed_cases_top7_outside_china)
plt_cum_confirmed_cases_by_country<-ggplot(confirmed_cases_top7_outside_china, aes(date, cum_cases, color = country)) +
  geom_line() +
  ylab("Cumulative confirmed cases")+
  ggtitle("Covid cases by country")
plt_cum_confirmed_cases_by_country
```


plots i want to show 
```{r}
plt_cum_confirmed_cases_worldwide<-ggplot(data = confirmed_cases_worldwide, aes(x=date, y=cum_cases))+geom_line()+ylab("Cumulative confirmed cases")+ ggtitle("Covid cases worldwide")
plt_cum_confirmed_cases_worldwide

plt_cum_confirmed_cases_china_vs_world <- ggplot(data=confirmed_cases_china_vs_world, aes(x= date, y= cum_cases, color = is_china)) +
  geom_line() +
  ylab("Cumulative confirmed cases") +
  ggtitle("Covid cases sepearted by China and non china countries")
plt_cum_confirmed_cases_china_vs_world

plt_not_china_trend_line<- ggplot(data = no_china, aes(x=date, y=cum_cases)) + geom_line() +ylab("cumulative confirmed cases")+ggtitle("Covid Cases in non-china countries")
plt_not_china_trend_line

plt_yes_china_trend_line<- ggplot(data = yes_china, aes(x=date, y=cum_cases)) + geom_line() +ylab("cumulative confirmed cases")+ggtitle("Covid Cases in china")+ylim(c(0,125000))
plt_yes_china_trend_line
```
Same plots with legend
```{r}
who_events <- tribble(
  ~ date, ~ event,
  "2020-01-30", "Global health\nemergency\ndeclared",
  "2020-03-11", "Pandemic\ndeclared",
  "2020-02-13", "China reporting\nchanged"
) %>%
  mutate(date = as.Date(date))

plt_cum_confirmed_cases_worldwide_legend<-ggplot(data = confirmed_cases_worldwide, aes(x=date, y=cum_cases))+geom_line()+ylab("Cumulative confirmed cases")+ ggtitle("Covid cases worldwide")+
  geom_vline(aes(xintercept = date), data = who_events, linetype = "dashed")+ geom_label(aes(date, label = event), data = who_events, y = 1e5, color="black")
plt_cum_confirmed_cases_worldwide_legend

plt_cum_confirmed_cases_china_vs_world_legend<-plt_cum_confirmed_cases_china_vs_world +
  geom_vline(aes(xintercept = date), data = who_events, linetype = "dashed") +
  geom_label(aes(date, label = event), data = who_events, y = 1e5, color="black")
plt_cum_confirmed_cases_china_vs_world_legend

plt_not_china_trend_line_legend<- ggplot(data = no_china, aes(x=date, y=cum_cases)) + geom_line() +ylab("cumulative confirmed cases")+ggtitle("Covid Cases in non-china countries")+
  geom_vline(aes(xintercept = date), data = who_events, linetype = "dashed") +
  geom_label(aes(date, label = event), data = who_events, y = 1e5, color="black")
plt_not_china_trend_line_legend

plt_yes_china_trend_line_legend<- ggplot(data = yes_china, aes(x=date, y=cum_cases)) + geom_line() +ylab("cumulative confirmed cases")+ggtitle("Covid Cases in china")+ylim(c(0,125000))+
  geom_vline(aes(xintercept = date), data = who_events, linetype = "dashed") +
  geom_label(aes(date, label = event), data = who_events, y = 1e5, color="black")
plt_yes_china_trend_line_legend
```




